---
title: "insomnia"
author: "Jane Han"
date: "2023-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r insomnia}
#install.packages("utils")
library(utils)

#setwd("/Users/mindmatter/Dropbox (Dartmouth College)/Courseworks/QBS181/Project_Insomnia")
data <- read.csv("./data/insomnia_data_cleaned.csv", header = TRUE)

setwd("/Users/freyam/Desktop/Dartmouth/Courses/002 Fall Term/001_Core Courses/QBS 181 - Data Wrangling /Final Group Project/Codes and cleaned data")
data <- read.csv("Insomnia_final.csv")

# data check 
#summary(data)
head(data, 30)
```

## Insomnia Data regression check 
check the regression between two columns 
columns: ISI_total - GCTI_total
analysis: pearson correlation 
```{r}
# Compute the Pearson correlation coefficient
correlation_coefficient <- cor(data$ISI_total, data$GCTI_total, method = "pearson")

# Print the result
print(correlation_coefficient)

correlation_test <- cor.test(data$ISI_total, data$GCTI_total)

# Print the results
print(correlation_test)
```

## Scatter Plot

```{r pressure, echo=FALSE}
# Scatter plot
plot(data$ISI_total, data$GCTI_total, main="Scatter plot of ISI_total vs GCTI_total",
     xlab="ISI_total", ylab="GCTI_total", col="blue", pch=16)

# Add a regression line to the plot
abline(lm(data$GCTI_total ~ data$ISI_total), col="red")
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Self Feedback for future analysis, notes
[_] could just make a heatmap of all columns as well?

```{r}
# Assuming 'data' is your data frame containing all the variables

# List of column names for which you want to calculate pairwise correlations
columns_to_analyze <- c("PDS_FEMALE", "PDS_MALE", "ISI_total", "PSQI_total", "BDI_total",
                        "ASHS_total", "ASHS_physiological", "ASHS_cognitive", "ASHS_emotional", 
                        "ASHS_SleepEnvirnmont", "ASHS_DaytimeSleep", "ASHS_substances", 
                        "ASHS_bedtimeRoutine", "ASHS_sleepStability", "ASHS_BedroomSharing", 
                        "DBAS_total", "FIRST_total", "GCTI_total", "GCTI_anxiety", "GCTI_reflection",
                        "GCTI_worries", "GCTI_thoughts", "GCTI_negativeAffect", "STAI_Y_total", 
                        "NEO_neuroticism", "NEO_extraversion", "NEO_openness", "NEO_agreeableness", 
                        "NEO_Conscientiousness", "MEQr_total", "PSRS_PrR", "PSRS_RWO", "PSRS_RSC", 
                        "PSRS_FRa", "PSRS_RSE", "PSRS_total", "PSS_total", "TCQI_R_Total", 
                        "TCQIR_Aggressive_supression", "TCQIR_cognitive_distraction", "TCQIR_reappraisal", 
                        "TCQIR_behavtioral_distraction", "TCQIR_social_avoidance", "TCQIR_worry", 
                        "ACE_tot", "asq_home", "asq_school", "asq_attendance", "asq_romantic", "asq_peer",
                        "asq_teacher", "asq_future", "asq_leisure", "asq_finance", "asq_responsibility",
                        "casq_total", "casq_sleepy", "casq_alert", "cope_disengage_su", "cope_growth",
                        "cope_disengage_mental", "cope_emotions", "cope_socialsupp_instr", "cope_active",
                        "cope_denial", "cope_religion", "cope_humor", "cope_disengage_emo", "cope_restraint",
                        "cope_socialsupp_emo", "cope_acccept", "cope_suppression", "cope_planning",
                        "ders_nonaccpetance", "ders_total", "ders_goals", "ders_impulse", "ders_awareness",
                        "ders_strategies", "ders_clarity")

# Create an empty list to store the results
correlation_results <- list()

# Generate all combinations of the column names
column_combinations <- combn(columns_to_analyze, 2)

# Function to calculate Pearson correlation for a pair
calculate_correlation <- function(column_pair) {
  col1 <- data[[column_pair[1]]]
  col2 <- data[[column_pair[2]]]
  complete_cases <- complete.cases(col1, col2)
  
  # Check if there are enough complete cases
  if (sum(complete_cases) > 2) {  # More than two observations are needed to calculate correlation
    result <- cor.test(col1[complete_cases], col2[complete_cases], method = "pearson")
    return(list(column_pair = column_pair, correlation_coefficient = result$estimate, 
                p_value = result$p.value, method = "pearson"))
  } else {
    return(list(column_pair = column_pair, correlation_coefficient = NA, 
                p_value = NA, method = "pearson", 
                warning = "Not enough finite observations"))
  }
}

# Apply the function to each pair
correlation_results <- apply(column_combinations, 2, calculate_correlation)

# Convert the results to a data frame, including a warning column for pairs with insufficient data
correlation_df <- do.call(rbind, lapply(correlation_results, function(x) {
  data.frame(column1 = x$column_pair[1], column2 = x$column_pair[2],
             correlation_coefficient = x$correlation_coefficient,
             p_value = x$p_value, warning = ifelse(is.na(x$correlation_coefficient), x$warning, ""))
}))

# View the dataframe
View(correlation_df)

```


# visualize with heatmap only when the p-value is significant (less than 0.05)

```{r}
# Assuming correlation_df is your dataframe with all correlation results
significant_correlations <- correlation_df[correlation_df$p_value < 0.05, ]

# Extract the unique column names that appear in the significant correlations
columns_in_significant_correlations <- unique(c(significant_correlations$column1, significant_correlations$column2))

# Create a blank matrix for the heatmap
correlation_matrix <- matrix(NA, 
                             nrow = length(columns_in_significant_correlations), 
                             ncol = length(columns_in_significant_correlations),
                             dimnames = list(columns_in_significant_correlations, columns_in_significant_correlations))

# Fill in the matrix with the significant correlation coefficients
for(i in 1:nrow(significant_correlations)) {
  row <- significant_correlations[i, ]
  col1 <- which(columns_in_significant_correlations == row$column1)
  col2 <- which(columns_in_significant_correlations == row$column2)
  correlation_matrix[col1, col2] <- row$correlation_coefficient
  correlation_matrix[col2, col1] <- row$correlation_coefficient  # For symmetry
}

# Since we only have significant correlations, we can replace NA with 0 for better visualization
correlation_matrix[is.na(correlation_matrix)] <- 0

# Plot the heatmap
library(ggplot2)

# Melt the matrix for ggplot
library(reshape2)
melted_correlation_matrix <- melt(correlation_matrix)

# Create the heatmap
ggplot(data = melted_correlation_matrix, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1, 1), space = "Lab", name="Pearson\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.title = element_blank())

ggsave("heatmap.png", plot = last_plot(), width = 12, height = 10, dpi = 300)

```



```{r}
#install.packages("plotly")
library(plotly)

# Convert the melted matrix for plotly
plotly_data <- as.data.frame(melted_correlation_matrix)

# Create the plotly interactive heatmap
plot_ly(x = plotly_data$Var2, y = plotly_data$Var1, z = plotly_data$value, 
        type = "heatmap", colors = colorRamp(c("blue", "white", "red")))
```
comment: 
Clustering - Sometimes clustering the rows and columns can lead to a more readable heatmap. Many heatmap functions have built-in clustering features that you can enable.

freya's comment

we have think of now choosing from these exploratory data/combination

figure out how to incoorporate population analysis 

```{r}
# Assuming 'data' is your dataframe and it has columns 'Asian' and 'White' along with the variable of interest, for example 'ISI_total'
# Create a new race factor variable where 1 represents 'White' and 2 represents 'Asian'
data$Race <- NA  # Create an empty column for race
data$Race[data$White == 1] <- 'White'
data$Race[data$Asian == 1] <- 'Asian'

# Convert the new Race column to a factor
data$Race <- factor(data$Race)

# Now, select only the rows where Race is 'Asian' or 'White'
data_subset <- data[data$Race %in% c('Asian', 'White'), ]

# Run a two-way ANOVA with Race and ISI_total (as an example)
# Replace 'ISI_total' with the variable of interest from your heatmap
result <- aov(ISI_total ~ Race , data = data_subset)
summary(result)


```

```{r}
# Assuming 'data' is your dataframe and it has columns 'Asian' and 'White' along with the variable of interest, for example 'ISI_total'
# Create a new race factor variable where 1 represents 'White' and 2 represents 'Asian'
data$Race <- NA  # Create an empty column for race
data$Race[data$White == 1] <- 'White'
data$Race[data$Asian == 1] <- 'Asian'

# Convert the new Race column to a factor
data$Race <- factor(data$Race)

# Now, select only the rows where Race is 'Asian' or 'White'
data_subset <- data[data$Race %in% c('Asian', 'White'), ]

# Run a two-way ANOVA with Race and ISI_total (as an example)
# Replace 'ISI_total' with the variable of interest from your heatmap
result_gcti_isi <- aov(ISI_total ~ Race + GCTI_total + Race:GCTI_total, data = data_subset)
summary(result_gcti_isi)

```
```{r}
# Run a two-way ANOVA with Race and ISI_total
result_isi_bdi <- aov(BDI_total ~ Race + ISI_total + Race:ISI_total, data = data_subset)
summary(result_isi_bdi)
```
```{r}
# Run a two-way ANOVA with Race and ISI_total
result_ashs_bdi <- aov(BDI_total ~ Race + ASHS_total + Race:ASHS_total, data = data_subset)
summary(result_ashs_bdi)
```


# White and Asian
with Insomnia data
```{r}
library(ggplot2)

# Create a new dataframe that contains the means of the dependent variable for each combination
# of levels of the independent variables.
# group_means <- aggregate(BDI_total ~ Race + ISI_total + ASHS_total, data = data, FUN = mean)

# Interaction plot for Race and ISI_total
ggplot(data_subset, aes(x = ISI_total, y = BDI_total, color = Race, group = Race)) +
  geom_smooth(aes(y = BDI_total), method = "lm", se = FALSE, color ="grey") +
  geom_point(aes(color = Race), alpha = 0.6) +
  theme_minimal() +
  labs(title = "Interaction of Race, Insomnia Severity, and Depression", x = "ISI_total", y = "BDI_total")

# Interaction plot for Race and ASHS_total
ggplot(data_subset, aes(x = ASHS_total, y = BDI_total, color = Race, group = Race)) +
  geom_smooth(aes(y = BDI_total), method = "lm", se = FALSE, color ="grey") +
  geom_point(aes(color = Race), alpha = 0.6) +
  theme_minimal() +
  labs(title = "Interaction of  Race, Sleep Hygiene, and Depression", x = "ASHS_total", y = "BDI_total")
```

# All race
with Insomnia_Final data
```{r}
library(ggplot2)

# Create a new dataframe that contains the means of the dependent variable for each combination
# of levels of the independent variables.
# group_means <- aggregate(BDI_total ~ Race + ISI_total + ASHS_total, data = data, FUN = mean)

# Interaction plot for Race and ISI_total
ggplot(Insomnia_Final, aes(x = ISI_total, y = BDI_total, color = Race, group = Race)) +
  geom_smooth(aes(y = BDI_total), method = "lm", se = FALSE, color ="grey") +
  geom_point(aes(color = Race), alpha = 0.6) +
  theme_minimal() +
  labs(title = "Interaction of Race and ISI_total", x = "ISI_total", y = "BDI_total")

# Interaction plot for Race and ASHS_total
ggplot(Insomnia_Final, aes(x = ASHS_total, y = BDI_total, color = Race, group = Race)) +
  geom_smooth(aes(y = BDI_total), method = "lm", se = FALSE, color ="grey") +
  geom_point(aes(color = Race), alpha = 0.6) +
  theme_minimal() +
  labs(title = "Interaction of Race and ASHS", x = "ASHS_total", y = "BDI_total")
```


```{r}
library(ggplot2)

# Calculate the means for each group
group_means <- aggregate(ISI_total ~ Race + GCTI_total, data = data_subset, mean)

# Convert the other_variable to a factor if it's not already
group_means$GCTI_total <- factor(group_means$GCTI_total)

# Plotting
ggplot(group_means, aes(x = GCTI_total, y = ISI_total, group = Race, color = Race)) +
  geom_line() +  # Add the lines for each Race
  geom_point() + # Add the points to indicate observed group means
  theme_minimal() +
  labs(x = "Mean of GCTI_total", y = "Mean of ISI_total", title = "Interaction Plot for Two-way ANOVA")

```



```{r}
library(ggplot2)

# Calculate the means for each group
group_means <- aggregate(ISI_total ~ Race + PSQI_total, data = data_subset, mean)

# Convert the other_variable to a factor if it's not already
group_means$PSQI_total <- factor(group_means$PSQI_total)

# Plotting
ggplot(group_means, aes(x = PSQI_total, y = ISI_total, group = Race, color = Race)) +
  geom_line() +  # Add the lines for each Race
  geom_point() + # Add the points to indicate observed group means
  theme_minimal() +
  labs(x = "Mean of PSQI_total", y = "Mean of ISI_total", title = "Interaction Plot for Two-way ANOVA")
```

```{r}
library(ggplot2)

# Calculate the means for each group
group_means <- aggregate(ISI_total ~ Race + PDS_MALE, data = data_subset, mean)

# Convert the other_variable to a factor if it's not already
group_means$PDS_MALE <- factor(group_means$PDS_MALE)

# Plotting
ggplot(group_means, aes(x = PDS_MALE, y = ISI_total, group = Race, color = Race)) +
  geom_line() +  # Add the lines for each Race
  geom_point() + # Add the points to indicate observed group means
  theme_minimal() +
  labs(x = "Mean of PDS_MALE", y = "Mean of ISI_total", title = "Interaction Plot for Two-way ANOVA")
```

```{r}
library(ggplot2)

# Calculate the means for each group
group_means <- aggregate(ISI_total ~ Race + SubGroup, data = data_subset, mean)

# Convert the other_variable to a factor if it's not already
group_means$SubGroup <- factor(group_means$SubGroup)

# Plotting
ggplot(group_means, aes(x = SubGroup, y = ISI_total, group = SubGroup, color = SubGroup)) +
  geom_line() +  # Add the lines for each Race
  geom_point() + # Add the points to indicate observed group means
  theme_minimal() +
  labs(x = "Mean of SubGroup", y = "Mean of ISI_total", title = "Plot for Two-way ANOVA")
```

```{r}
library(ggplot2)

# Calculate the means for each group
group_means <- aggregate(ISI_total ~ Race + Group, data = data_subset, mean)

# Convert the other_variable to a factor if it's not already
group_means$SubGroup <- factor(group_means$Group)

# Plotting
ggplot(group_means, aes(x = Group, y = ISI_total, group = Group, color = Group)) +
  geom_line() +  # Add the lines for each Race
  geom_point() + # Add the points to indicate observed group means
  theme_minimal() +
  labs(x = "Mean of Group", y = "Mean of ISI_total", title = "Plot for Two-way ANOVA")
```
```{r}
library(ggplot2)

# Create a new dataframe that contains the means of the dependent variable for each combination
# of levels of the independent variables.
group_means <- aggregate(BDI_total ~ Race + ISI_total + ASHS_total, data = data, FUN = mean)

# Interaction plot for Race and ISI_total
ggplot(group_means, aes(x = ISI_total, y = BDI_total, color = Race, group = Race)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(title = "Interaction of Race and ISI_total", x = "ISI_total", y = "BDI_total")

# Interaction plot for Race and ASHS_total
ggplot(group_means, aes(x = ASHS_total, y = BDI_total, color = Race, group = Race)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(title = "Interaction of Race, ASHS, and BDI", x = "ASHS_total", y = "BDI_total")

```


```{r}
library(gridExtra)
library(ggplot2)

# Define the variables and titles
variables <- c("ASHS_total")
titles <- c("Depression")

p <- ggplot(Insomnia_Final, aes(x = ASHS_total, y = BDI_total, color = Race)) +
  geom_point(aes(color = Race), alpha = 0.6) +
  geom_smooth(aes(y = BDI_total), method = "lm", se = FALSE, color = "steelblue") +
  labs(title = paste(title_prefix, "vs", x_var)) +
  scale_color_brewer(palette = "RdYlGn") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 10), 
    axis.title = element_text(size = 10), 
    legend.position = "right", 
    legend.background = element_rect(fill = "white", colour = "grey50"),
    legend.text = element_text(size = 5), 
    legend.title = element_text(face = "bold", size = 7), 
    legend.key.size = unit(0.5, "lines") 
  )

# Print the scatter plot
print(p)

# Relationship between Age, BDI, and ASHS Scores with Race color distinction
bdi_ashs_scatter <- ggplot(Insomnia_Final, aes(x = ASHS_total, y = BDI_total, color = Race)) +
  geom_point(aes(size = 0.5), alpha = 0.6) +
  geom_smooth(aes(y = BDI_total), method = "lm", se = FALSE, color = "black") +
  labs(title = "Relationship between Race, ASHS Scores, and BDI") +
  scale_color_brewer(palette = "RdYlGn") + 
  base_theme +
  theme(legend.title = element_text(size = 8, face = "bold"),
        legend.text = element_text(size = 8))

# Print the scatter plot
print(bdi_ashs_scatter)
```




```{r}
library(ggplot2)

# Create a new dataframe that contains the means of the dependent variable for each combination
# of levels of the independent variables.
group_means <- aggregate(BDI_total ~ Race + ISI_total + ASHS_total, data = data, FUN = mean)

# Interaction plot for ISI_total and ASHS_total
ggplot(group_means, aes(x = ISI_total, y = BDI_total, color = ASHS_total, group = ASHS_total)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(title = "Interaction of ISI, ASHS, and BDI", x = "ISI_total", y = "BDI_total")

```









## 3-way 

ANOVA analysis 

```{r}
# Assuming 'data' is your data frame, and it contains 'race', 'group', and 'ISI_total' columns
# 'race' and 'group' should be converted to factors if they are not already

data$Race <- factor(data$Race)

# Perform the three-way ANOVA
three_way_anova_result <- aov(BDI_total ~ Race * ISI_total * ASHS_total, data = data)

# Get the summary of the ANOVA
summary(three_way_anova_result)
```
# Visualizations
```{r}
library(ggplot2)
library(gridExtra)
library(RColorBrewer)

# Set the base theme for our plots
base_theme <- theme_minimal() +
              theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 14, family = "Arial"),
                    plot.subtitle = element_text(hjust = 0.5))
color_palette <- scale_fill_brewer(palette = "RdYlGn") 

# BDI Score Distribution by Race
bdi_race_boxplot <- ggplot(Insomnia_Final, aes(x = Race, y = BDI_total, fill = Race)) +
  geom_boxplot() +
  labs(title = "BDI Score Distribution by Race") +
  color_palette +
  base_theme +
  theme(legend.position = "none")

# ASHS Score Distribution by Race
ashs_race_boxplot <- ggplot(Insomnia_Final, aes(x = Race, y = ASHS_total, fill = Race)) +
  geom_boxplot() +
  labs(title = "ASHS Score Distribution by Race") +
  color_palette +
  base_theme +
  theme(legend.position = "none")

# Relationship between Age, BDI, and ASHS Scores with Race color distinction
bdi_ashs_scatter <- ggplot(Insomnia_Final, aes(x = ASHS_total, y = BDI_total, color = Race)) +
  geom_point(aes(size = 0.5), alpha = 0.6) +
  geom_smooth(aes(y = BDI_total), method = "lm", se = FALSE, color = "black") +
  labs(title = "Relationship between Race, ASHS Scores, and BDI") +
  scale_color_brewer(palette = "RdYlGn") + 
  base_theme +
  theme(legend.title = element_text(size = 8, face = "bold"),
        legend.text = element_text(size = 8))

# Arrange the boxplots in a grid for better comparison
grid.arrange(bdi_race_boxplot, ashs_race_boxplot, nrow = 2)

# Print the scatter plot
print(bdi_ashs_scatter)

```

## 4-way 

```{r}
library(ggplot2)
library(gridExtra)
library(RColorBrewer)

# Set the base theme for our plots
base_theme <- theme_minimal() +
              theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 14, family = "Arial"),
                    plot.subtitle = element_text(hjust = 0.5))
color_palette <- scale_fill_brewer(palette = "RdYlGn") 

# BDI Score Distribution by Race
bdi_race_boxplot <- ggplot(Insomnia_Final, aes(x = Race, y = BDI_total, fill = Race)) +
  geom_boxplot() +
  labs(title = "BDI Score Distribution by Race") +
  color_palette +
  base_theme +
  theme(legend.position = "none")

# ASHS Score Distribution by Race
ashs_race_boxplot <- ggplot(Insomnia_Final, aes(x = Race, y = ASHS_total, fill = Race)) +
  geom_boxplot() +
  labs(title = "ASHS Score Distribution by Race") +
  color_palette +
  base_theme +
  theme(legend.position = "none")

# Relationship between Age, BDI, and ASHS Scores with Race color distinction
age_bdi_ashs_scatter <- ggplot(Insomnia_Final, aes(x = Age, y = BDI_total, color = Race)) +
  geom_point(aes(size = ASHS_total), alpha = 0.6) +
  geom_smooth(aes(y = BDI_total), method = "lm", se = FALSE, color = "black") +
  labs(title = "Relationship between Age, Race, ASHS Scores, and BDI") +
  scale_color_brewer(palette = "RdYlGn") + 
  base_theme +
  theme(legend.title = element_text(size = 8, face = "bold"),
        legend.text = element_text(size = 8))

# Arrange the boxplots in a grid for better comparison
grid.arrange(bdi_race_boxplot, ashs_race_boxplot, nrow = 2)

# Print the scatter plot
print(age_bdi_ashs_scatter)

```

```{r}

```

```{r}

```
