---
title: "insomnia"
author: "Jane Han"
date: "2023-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r insomnia}
#install.packages("utils")
library(utils)

setwd("/Users/mindmatter/Dropbox (Dartmouth College)/Courseworks/QBS181/Project_Insomnia")
data <- read.csv("./data/insomnia_data_cleaned.csv", header = TRUE)

# data check 
#summary(data)
head(data, 30)
```

## Insomnia Data regression check 
check the regression between two columns 
columns: ISI_total - GCTI_total
analysis: pearson correlation 
```{r}
# Compute the Pearson correlation coefficient
correlation_coefficient <- cor(data$ISI_total, data$GCTI_total, method = "pearson")

# Print the result
print(correlation_coefficient)

correlation_test <- cor.test(data$ISI_total, data$GCTI_total)

# Print the results
print(correlation_test)
```

## Scatter Plot

```{r pressure, echo=FALSE}
# Scatter plot
plot(data$ISI_total, data$GCTI_total, main="Scatter plot of ISI_total vs GCTI_total",
     xlab="ISI_total", ylab="GCTI_total", col="blue", pch=16)

# Add a regression line to the plot
abline(lm(data$GCTI_total ~ data$ISI_total), col="red")
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Self Feedback for future analysis, notes
[_] could just make a heatmap of all columns as well?

```{r}
# Assuming 'data' is your data frame containing all the variables

# List of column names for which you want to calculate pairwise correlations
columns_to_analyze <- c("PDS_FEMALE", "PDS_MALE", "ISI_total", "PSQI_total", "BDI_total",
                        "ASHS_total", "ASHS_physiological", "ASHS_cognitive", "ASHS_emotional", 
                        "ASHS_SleepEnvirnmont", "ASHS_DaytimeSleep", "ASHS_substances", 
                        "ASHS_bedtimeRoutine", "ASHS_sleepStability", "ASHS_BedroomSharing", 
                        "DBAS_total", "FIRST_total", "GCTI_total", "GCTI_anxiety", "GCTI_reflection",
                        "GCTI_worries", "GCTI_thoughts", "GCTI_negativeAffect", "STAI_Y_total", 
                        "NEO_neuroticism", "NEO_extraversion", "NEO_openness", "NEO_agreeableness", 
                        "NEO_Conscientiousness", "MEQr_total", "PSRS_PrR", "PSRS_RWO", "PSRS_RSC", 
                        "PSRS_FRa", "PSRS_RSE", "PSRS_total", "PSS_total", "TCQI_R_Total", 
                        "TCQIR_Aggressive_supression", "TCQIR_cognitive_distraction", "TCQIR_reappraisal", 
                        "TCQIR_behavtioral_distraction", "TCQIR_social_avoidance", "TCQIR_worry", 
                        "ACE_tot", "asq_home", "asq_school", "asq_attendance", "asq_romantic", "asq_peer",
                        "asq_teacher", "asq_future", "asq_leisure", "asq_finance", "asq_responsibility",
                        "casq_total", "casq_sleepy", "casq_alert", "cope_disengage_su", "cope_growth",
                        "cope_disengage_mental", "cope_emotions", "cope_socialsupp_instr", "cope_active",
                        "cope_denial", "cope_religion", "cope_humor", "cope_disengage_emo", "cope_restraint",
                        "cope_socialsupp_emo", "cope_acccept", "cope_suppression", "cope_planning",
                        "ders_nonaccpetance", "ders_total", "ders_goals", "ders_impulse", "ders_awareness",
                        "ders_strategies", "ders_clarity")

# Create an empty list to store the results
correlation_results <- list()

# Generate all combinations of the column names
column_combinations <- combn(columns_to_analyze, 2)

# Function to calculate Pearson correlation for a pair
calculate_correlation <- function(column_pair) {
  col1 <- data[[column_pair[1]]]
  col2 <- data[[column_pair[2]]]
  complete_cases <- complete.cases(col1, col2)
  
  # Check if there are enough complete cases
  if (sum(complete_cases) > 2) {  # More than two observations are needed to calculate correlation
    result <- cor.test(col1[complete_cases], col2[complete_cases], method = "pearson")
    return(list(column_pair = column_pair, correlation_coefficient = result$estimate, 
                p_value = result$p.value, method = "pearson"))
  } else {
    return(list(column_pair = column_pair, correlation_coefficient = NA, 
                p_value = NA, method = "pearson", 
                warning = "Not enough finite observations"))
  }
}

# Apply the function to each pair
correlation_results <- apply(column_combinations, 2, calculate_correlation)

# Convert the results to a data frame, including a warning column for pairs with insufficient data
correlation_df <- do.call(rbind, lapply(correlation_results, function(x) {
  data.frame(column1 = x$column_pair[1], column2 = x$column_pair[2],
             correlation_coefficient = x$correlation_coefficient,
             p_value = x$p_value, warning = ifelse(is.na(x$correlation_coefficient), x$warning, ""))
}))

# View the dataframe
View(correlation_df)

```


# visualize with heatmap only when the p-value is significant (less than 0.05)

```{r}
# Assuming correlation_df is your dataframe with all correlation results
significant_correlations <- correlation_df[correlation_df$p_value < 0.05, ]

# Extract the unique column names that appear in the significant correlations
columns_in_significant_correlations <- unique(c(significant_correlations$column1, significant_correlations$column2))

# Create a blank matrix for the heatmap
correlation_matrix <- matrix(NA, 
                             nrow = length(columns_in_significant_correlations), 
                             ncol = length(columns_in_significant_correlations),
                             dimnames = list(columns_in_significant_correlations, columns_in_significant_correlations))

# Fill in the matrix with the significant correlation coefficients
for(i in 1:nrow(significant_correlations)) {
  row <- significant_correlations[i, ]
  col1 <- which(columns_in_significant_correlations == row$column1)
  col2 <- which(columns_in_significant_correlations == row$column2)
  correlation_matrix[col1, col2] <- row$correlation_coefficient
  correlation_matrix[col2, col1] <- row$correlation_coefficient  # For symmetry
}

# Since we only have significant correlations, we can replace NA with 0 for better visualization
correlation_matrix[is.na(correlation_matrix)] <- 0

# Plot the heatmap
library(ggplot2)

# Melt the matrix for ggplot
library(reshape2)
melted_correlation_matrix <- melt(correlation_matrix)

# Create the heatmap
ggplot(data = melted_correlation_matrix, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1, 1), space = "Lab", name="Pearson\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        axis.title = element_blank())

ggsave("heatmap.png", plot = last_plot(), width = 12, height = 10, dpi = 300)

```



```{r}
#install.packages("plotly")
library(plotly)

# Convert the melted matrix for plotly
plotly_data <- as.data.frame(melted_correlation_matrix)

# Create the plotly interactive heatmap
plot_ly(x = plotly_data$Var2, y = plotly_data$Var1, z = plotly_data$value, 
        type = "heatmap", colors = colorRamp(c("blue", "white", "red")))
```
comment: 
Clustering - Sometimes clustering the rows and columns can lead to a more readable heatmap. Many heatmap functions have built-in clustering features that you can enable.


```{r}

```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```

```{r}
```


