---
title: "181_Final Group Project"
Author: Anything Health (Group Insomnia)
output: pdf_document
date: "2023-10-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preparation ##
*Please note that previous data cleaning and wrangling processes have also been performed in Excel prior to this R analysis*

```{r}

# Set working directory 
setwd("/Users/freyam/Desktop/Dartmouth/Courses/002 Fall Term/001_Core Courses/QBS 181 - Data Wrangling /Final Group Project/Codes and cleaned data")

# Read in the data
Insomnia <- read.csv("insomnia_data_cleaned.csv")

# Dataset inspection
head(Insomnia)
colnames(Insomnia)

```

## Data Cleaning ##

```{r}
library(mice) 

# Impute missing data using the mice package
mice_mod <- mice(Insomnia, m = 1, method = 'pmm', maxit = 5)
Insomnia_clean <- complete(mice_mod)

# Remove rows with any missing value if still present
Insomnia_clean <- Insomnia_clean %>% 
  drop_na()

# Save the cleaned dataset
write.csv(Insomnia_clean, "Insomnia_final.csv", row.names = FALSE)

```

```{r}
# Read in the final dataset again
Insomnia_Final <- read.csv("Insomnia_final.csv")

# Dataset inspection
head(Insomnia_Final)

```

## Preliminary Step No.1: Compare Age and Race by subgroups ##
There are 3 different sub-groups in this dataset:
- 0: Control
- 1: Clean Insomnia
- 2: Sub-clinical Insomnia

```{r}

library(tidyverse)
library(ggplot2)
library(dplyr)
library(viridis)

# Combine individual race columns into one for the Insomnia_Final dataframe
Insomnia_Final <- Insomnia_Final %>%
  mutate(
    Race = case_when(
      American_Indian == 1 ~ "American_Indian",
      Asian == 1 ~ "Asian",
      Native_Hawaiian == 1 ~ "Native_Hawaiian",
      Black == 1 ~ "Black",
      White == 1 ~ "White",
      TRUE ~ "Other"
    )
  )

# Density plot for age distribution within each race 
densityplot_age_by_race <- ggplot(Insomnia_Final, aes(x = Age, fill = Race)) +
  geom_density(alpha = 0.7) + 
  scale_fill_viridis(discrete = TRUE) + 
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"), 
    legend.position = "bottom" 
  ) +
  labs(
    title = "Age Distribution by Race",
    x = "Age",
    y = "Density"
  )

# Bar plot for subgroup distribution by race 
barplot_group_by_race <- ggplot(Insomnia_Final, aes(x = SubGroup, fill = Race)) +
  geom_bar(position = position_dodge()) + 
  scale_fill_viridis(discrete = TRUE) + 
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"), 
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1) 
  ) +
  labs(
    title = "SubGroup Distribution by Race",
    x = "SubGroup",
    y = "Count"
  )

# Render the plots
print(densityplot_age_by_race)
print(barplot_group_by_race)

```

We can now take a look at the two most important variables (e.g., "ISI_total: Insomnia severity Index" and "BDI_total: Becks Depression Inventory") across ages and races to have a better understanding of our data structure.

```{r}

library(tidyverse)
library(ggplot2)
library(viridis)
library(gridExtra)


# Create a long format dataframe to include both ISI_total and BDI_total
long_data <- Insomnia_Final %>%
  select(Age, Race, ISI_total, BDI_total) %>%
  gather(key = "ScoreType", value = "Score", -Age, -Race)

# Scatterplot for both ISI_total and BDI_total vs Age, faceted by Race and ScoreType
scatterplot_age_scores <- ggplot(long_data, aes(x = Age, y = Score, color = Race)) +
  geom_point(alpha = 0.6, size = 3) +
  facet_grid(ScoreType ~ Race, scales = "free_y") +
  scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"), 
    legend.position = "bottom"
  ) +
  labs(
    title = "Total Scores by Age and Race",
    x = "Age",
    y = "Total Score"
  )

# Render the plot
print(scatterplot_age_scores)

```

Now, we will move on to perform some battery tests to compare and test correlations to see what variables are suitable for the analysis. Please note that all variables are randomly chosen from the dataset and not all variables are included in the combinations.

```{r}

# List of variable combinations 
combinations <- list(
  c("GCTI_total", "NEO_agreeableness"),
  c("ASHS_total", "NEO_neuroticism"),
  c("ISI_total", "NEO_extraversion"),
  c("PSQI_total", "NEO_openness"),
  c("BDI_total", "MEQr_total"),
  c("ASHS_physiological", "PSRS_total"),
  c("ASHS_cognitive", "NEO_Conscientiousness"),
  c("ASHS_emotional", "PSS_total"),
  c("ASHS_SleepEnvirnmont", "TCQI_R_Total"),
  c("ASHS_DaytimeSleep", "TCQIR_cognitive_distraction"),
  c("ASHS_substances", "NEO_agreeableness"),
  c("ASHS_bedtimeRoutine", "ZNEO_neuroticism"),
  c("ASHS_sleepStability", "ZNEO_openness"),
  c("ASHS_BedroomSharing", "ZPSRS_RSE"),
  c("GCTI_anxiety", "ZTCQIR_reappraisal"),
  c("GCTI_reflection", "ZSTAI_Y_total"),
  c("GCTI_worries", "Zders_awareness"),
  c("GCTI_thoughts", "ZDERS_total")
)

# Function to plot each combination
plot_combination <- function(vars) {
  ggplot(Insomnia, aes_string(x=vars[1], y=vars[2])) + 
    geom_point(aes(color=Race), alpha=0.6) +
    geom_smooth(method="lm", se=FALSE, color="black") +
    labs(title=paste("Scatter Plot of", vars[1], "and", vars[2]), 
         x=vars[1], 
         y=vars[2]) +
    theme_minimal()
}

# Generate and display plots
lapply(combinations, plot_combination)

```
Based on our the above battery tests and our defined data dictionaries, we have chosen the following variables for further investigation: 

- ISI_total: Insomnia severity Index
- BDI_total: Becks Depression Inventory
- GCTI_total: The Glasgow Content of Thoughts Inventory (e.g., anxiety, reflection, worries, thoughts, negativeAffect)
- ASHS_total: Adolescent Sleep Hygiene Scale (e.g., physiological, emotional, SleepEnvironment, Substances, bedtimeRoutine)


```{r}

library(GGally)
library(ggplot2)

# Define the columns of interest
cols_of_interest <- c("ISI_total", "BDI_total", "GCTI_total", "ASHS_total")

# Create the scatter plot matrix with a specified color
scatter_plot_matrix <- ggpairs(
  Insomnia_Final[, cols_of_interest],
  lower = list(
    continuous = wrap("points", color = "steelblue", alpha = 0.5) 
  ),
  upper = list(
    continuous = wrap("cor", color = "darkred")  
  ),
  diag = list(
    continuous = wrap("barDiag", fill = "#132B43") 
  )
)

# Print the plot
print(scatter_plot_matrix)

```

After investigating the total scores, we can now dig deeper into individual psychological factors and how they affect the severity of insomnia and depression. 

For example, since we are particularly interested in anxiety, worries, sleep stability, and the use of substances, we will examine the following four variables in relationship with insomnia and depression severity.

```{r}
library(gridExtra)
library(ggplot2)

# Define the variables and titles
variables <- c("ASHS_sleepStability", "ASHS_substances", "GCTI_anxiety", "GCTI_worries")
titles <- c("Insomnia Severity", "Insomnia Severity", "Depression", "Depression")

# Define a function to plot different plots into one grid
plot_variable <- function(df, x_var, y_var, title_prefix) {
  p <- ggplot(df, aes_string(x = x_var, y = y_var)) +
    geom_point(aes_string(color = x_var), alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE, color = "steelblue") +
    labs(title = paste(title_prefix, "vs", x_var),
         x = x_var,
         y = y_var) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 10), 
      axis.title = element_text(size = 10), 
      legend.position = "right", 
      legend.background = element_rect(fill = "white", colour = "grey50"),
      legend.text = element_text(size = 5), 
      legend.title = element_text(face = "bold", size = 7), 
      legend.key.size = unit(0.5, "lines") 
    )
  return(p)
}

# Create a list to hold plots
plot_list <- list()

# Generate plots
for (i in 1:length(variables)) {
  if (titles[i] == "Insomnia Severity") {
    plot_list[[i]] <- plot_variable(Insomnia_Final, variables[i], "ISI_total", titles[i])
  } else {
    plot_list[[i]] <- plot_variable(Insomnia_Final, variables[i], "BDI_total", titles[i])
  }
}

# Arrange the plots in a grid
do.call(gridExtra::grid.arrange, c(plot_list, ncol = 2))

```

Aim #1 (1.1) Part A

We can first take a look at GCTI and ASHS score distributions for different ages and races.

```{r}
library(ggplot2)
library(gridExtra)
library(RColorBrewer)

# Set the base theme for our plots
base_theme <- theme_minimal() +
              theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 14, family = "Arial"),
                    plot.subtitle = element_text(hjust = 0.5))
color_palette <- scale_fill_brewer(palette = "Blues") 

# GCTI Score Distribution by Race
gcti_race_boxplot <- ggplot(Insomnia_Final, aes(x = Race, y = GCTI_total, fill = Race)) +
  geom_boxplot() +
  labs(title = "GCTI Score Distribution by Race") +
  color_palette +
  base_theme +
  theme(legend.position = "none")

# ASHS Score Distribution by Race
ashs_race_boxplot <- ggplot(Insomnia_Final, aes(x = Race, y = ASHS_total, fill = Race)) +
  geom_boxplot() +
  labs(title = "ASHS Score Distribution by Race") +
  color_palette +
  base_theme +
  theme(legend.position = "none")

# Relationship between Age, GCTI, and ASHS Scores with Race color distinction
age_gcti_ashs_scatter <- ggplot(Insomnia_Final, aes(x = Age, y = GCTI_total, color = Race)) +
  geom_point(aes(size = ASHS_total), alpha = 0.6) +
  geom_smooth(aes(y = GCTI_total), method = "lm", se = FALSE, color = "black") +
  labs(title = "Relationship between Age, GCTI, and ASHS Scores") +
  scale_color_brewer(palette = "Blues") + 
  base_theme +
  theme(legend.title = element_text(size = 8, face = "bold"),
        legend.text = element_text(size = 8))

# Arrange the boxplots in a grid for better comparison
grid.arrange(gcti_race_boxplot, ashs_race_boxplot, nrow = 2)

# Print the scatter plot
print(age_gcti_ashs_scatter)

```

Aim #1 (1.1) Part B

We can now take a look at GCTI and ASHS score distributions for different subgroups

```{r}
library(ggplot2)
library(gridExtra)

# Create boxplots for GCTI distributions
gcti_boxplot <- ggplot(Insomnia_Final, aes(x = factor(SubGroup), y = GCTI_total, fill = factor(SubGroup))) +
  geom_boxplot() +
  scale_fill_manual(values = c("0" = "#1f77b4", "1" = "#aec7e8")) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "none"
  ) +
  labs(
    title = "GCTI Score Distribution by Subgroup",
    x = "Subgroup",
    y = "GCTI Score"
  )

# Create boxplots for ASHS distributions
ashs_boxplot <- ggplot(Insomnia_Final, aes(x = factor(SubGroup), y = ASHS_total, fill = factor(SubGroup))) +
  geom_boxplot() +
  scale_fill_manual(values = c("0" = "#1f77b4", "1" = "#aec7e8")) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "none"
  ) +
  labs(
    title = "ASHS Score Distribution by Subgroup",
    x = "Subgroup",
    y = "ASHS Score"
  )

# Combine the plots into a grid
combined_plots <- grid.arrange(gcti_boxplot, ashs_boxplot, nrow = 2)

# Print the combined plot grid
print(combined_plots)

```

Aim #1 (1.2) Part A

```{r}

library(ggplot2)
library(reshape2)
library(dplyr)

# Select variables starting with GCTI and ASHS along with ISI_total and BDI_total for the correlation matrix
selected_vars <- Insomnia_Final %>%
  select(ISI_total, BDI_total, starts_with("GCTI"), starts_with("ASHS"))

# Compute correlation matrix
cor_matrix <- cor(selected_vars, use = "complete.obs")

# Melt the correlation matrix for ggplot2
cor_data <- melt(cor_matrix)

# Define specific colors for blue-white-red gradient
blue_white_red_colors <- c("#132B43", "steelblue", "white", "indianred1", "#B2182B")

# Visualize the correlation matrix
correlation_matrix_heatmap <- ggplot(data = cor_data, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "white") + 
  geom_text(aes(label = sprintf("%.2f", value)), vjust = 1, color = "black", size = 2.5) + 
  scale_fill_gradientn(colors = blue_white_red_colors,
                       limits = c(-1, 1),
                       breaks = c(-1, -0.5, 0, 0.5, 1),
                       labels = c("-1", "-0.5", "0", "0.5", "1"),
                       name="Pearson\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        axis.text.y = element_text(angle = 45, vjust = 1, hjust=1),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  labs(title = "Correlation Matrix Heatmap") +
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1.5))

# Print the plot
print(correlation_matrix_heatmap)

```

Aim #1 (1.2) Part B (POTENTIALLY DELETE)

```{r}

library(ggplot2)
library(dplyr)
library(ggcorrplot)
library(RColorBrewer)

# Select only numerical variables for the correlation matrix
selected_vars <- Insomnia_Final %>% 
  select(Age, ISI_total, BDI_total, GCTI_total, ASHS_total)

# Create a correlation matrix
cor_mat <- cor(selected_vars, use = "complete.obs")

# Define a blue color palette
blue_palette <- colorRampPalette(c("#DCEBF7", "#6E9ECF", "#132B43"))(5)

# Generate a heatmap of the correlation matrix 
ggcorrplot(cor_mat, 
           method = "circle", 
           type = "lower", 
           lab = TRUE, 
           colors = blue_palette, 
           lab_size = 3.5, 
           tl.cex = 12,
           tl.col = "black", 
           tl.srt = 45, 
           title = "Correlation matrix of selected variables",
           ggtheme = theme_minimal(),
           outline.color = "black"
           ) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5)) + 
  labs(fill = "Correlation")


```

Aim #2 (2.1)

```{r}

library(lme4) 

# Predicting mental health from sleep quality using a linear model
model <- lm(BDI_total ~ PSQI_total + Age + Sex + Race, data = Insomnia_Final)
summary(model)

```

Aim #2 (2.2)

```{r}

library(lme4) 
library(interactions)

# Adding interaction between PSQI_total and Race in the model
interaction_model <- lm(BDI_total ~ PSQI_total * Race, data = Insomnia_Final)
summary(interaction_model)


```

Shiny Dashboard (Optional)

```{r}

# Loading necessary libraries
library(shiny)
library(shinydashboard)

# Building the Shiny dashboard
ui <- dashboardPage(
  dashboardHeader(title = "Insomnia Analysis Dashboard"),
  dashboardSidebar(
    sidebarMenu(
      menuItem("Subgroup Analysis", tabName = "subgroup_analysis", icon = icon("user")),
      menuItem("Regression Analysis", tabName = "regression_analysis", icon = icon("chart-line"))
    )
  ),
  dashboardBody(
    tabItems(
      tabItem(tabName = "subgroup_analysis",
              fluidRow(
                box(plotOutput("plotAgeRace"), width = 6),
                box(plotOutput("plotHeatmap"), width = 6)
              )),
      tabItem(tabName = "regression_analysis",
              fluidRow(
                box(plotOutput("plotRegression"), width = 6),
                box(plotOutput("plotInteraction"), width = 6)
              ))
    )
  )
)

server <- function(input, output) {
  output$plotAgeRace <- renderPlot({ plot_age_race })
  output$plotHeatmap <- renderPlot({ plot_heatmap })
  output$}

```

